---
title: "Identify Sepsis"
author: "Fei Ding"
date: "November 20, 2019"
output:
  html_document:
    df_print: paged
---




```{r setup, include = FALSE}
library(tidyverse)
library(scales)
library(kableExtra)
library(VennDiagram)
library(magrittr)

knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)
dob_dis <- read_csv("../output/dob_discrepancy.csv") 
cam_visits <- read_csv("../output/cam_stay_20190925.csv") 
sepsis3 <- read_csv("../output/sepsis3_20191014.csv")
rhee_sepsis <- read_csv("../output/sepsis_rhee_20191115.csv")
sepsis_comp <- cam_visits %>% 
  select(grid, adm_id) %>% 
  left_join(sepsis3 %>% 
              select(grid, adm_id, sofa, sepsis3, data_type)) %>% 
  left_join(rhee_sepsis %>% 
              select(grid, adm_id, rhee)) %>%  
  mutate(sepsis3_infection = if_else(is.na(sepsis3), 0, 1),
         sepsis3 = if_else(is.na(sepsis3), 0, sepsis3),
         rhee_infection = if_else(is.na(rhee), 0, 1),
         rhee = if_else(is.na(rhee), 0, rhee))

sepsis_comp %>% 
  semi_join(dob_dis)
sepsis_comp %>% 
  semi_join(dob_dis) %>% 
  distinct(grid)
```

# Update from last report
1. Rhee definition
    * correct platelet typo, the x$10^3$ was left out in Rhee's paper
    * Add phenylephrine and vasopressin and remove dobutamine
2. Lab value check
3. Patient Location

# DOB discrepancy
As discussed in last meeting, the following 15 patients with DOB discrepancy > 10 days were removed from our final results.  
```{r}
dob_dis %>% filter(abs(diff) > 10) #24 pts total, 15 had diff > 10 
cam_visits %<>% anti_join(dob_dis, by = "grid")
sepsis3 %<>% anti_join(dob_dis, by = "grid")
rhee_sepsis %<>% anti_join(dob_dis, by = "grid")
sepsis_comp %<>% anti_join(dob_dis, by = "grid")
```

# Sepsis-3 definition

**NOTE:** Deviation from original definition were **bold (need more discussion)** or _italic (decided deviation)_.

1. For each CAM-ICU encounter, identify the first episode of suspected infection _within $\pm1$ day of hopital admission date_.
    * If the antibiotic was given first, the culture sampling must have been obtained within 1 day.
    * If the culture sampling was first, the antibiotic must have been ordered within 3 days.
    * The onset date of infection was the date on which the first of these two events occured.
    * __We used only blood culutre instead of body fluid cultures (blood, urine, cerebrospinal fluid, etc)__.
2. For each suspected infection, find worst daily SOFA score for the time window from 2 day before to 1 day after the onset date of infection.
    * Central nervous system SOFA score calculated _based on RASS Method C in the Modified SOFA score paper_.  
    * Respiration SOFA score calculated using both PaO~2~/FiO~2~(PF) ratio _and SpO~2~/FiO~2~(SF) ratio_.
        * FiO~2~ values > 100 were excluded. 1 < FiO~2~ <= 100 were converted to 0-1. 
        * SpO~2~ values > 100 were converted to 100.
    * __We used only creatinine data for renal SOFA, no urine output data__.
    * __For vasopressor, there is a footnote "Catecholamine doses are given as $\mu g/kg/min$ for at least 1 hour", do we need to limit vasopressor to IV only (except dobutamine)?__
    * Note
        * All the lab/drug dates are from 3 days before admission date to discharge date. 
        * All non-numerical lab values were converted to upper/lower limit or mean of the range.  
   
    
|SOFA score|PF or SF ratio and Ventilation | Platelets, x$10^3/\mu L$| Bilirubin, mg/dL| Cardiovascular| Rass score | Creatinine, mg/dL |
--|-------------------------------------|----------|-----------|----------------------------------|--------------|---|----|
0 | PF $\ge$ 400 or SF $\ge$ 512          | $\ge$150 | <1.2     | MAP $\ge$70 mmHg                  |$\ge$ 0   | <1.2    |  
1 | PF < 400 or SF < 512                  | <150     | 1.2-1.9  | MAP < 70 mmHg                     | -1       | 1.2-1.9 |
2 | PF < 300 or SF < 357                  | <100     | 2.0-5.9  | any dopamine or dubutamine        | -2       | 2.0-3.4 |
3 | PF < 200 or SF < 214 with ventilation | <50      | 6.0-11.9 | any epinephrine or norepinephrine | -3       | 3.5-4.9 |
4 | PF < 100 or SF < 89 with ventilation  | <20      | >12.0    |                                   | -4 to -5 | >5.0    |

3. Suspected infection with a SOFA score $\ge2$ is considered sepsis __assuming baseline SOFA score is zero__.

## Distribution of SOFA score
`r paste0(comma(sum(sepsis3$sepsis3 == 1)), " (", percent(sum(sepsis3$sepsis3 == 1)/nrow(sepsis3)), ")")` out of the `r comma(nrow(sepsis3))` suspected infections had a SOFA score $\ge 2$.
```{r}
summary(sepsis3$sofa) 
ggplot(sepsis3) +
  geom_histogram(aes(x = sofa), binwidth = 1) +
  labs(title = "Distribution of SOFA score") 
```

## Missing Data
Missing data analysis below is based on the first worst SOFA of each suspected infection.  
Respiration system had the most missing data despite that we added SF ratio.  PaO2 had the most amount of data, which only covers fewer than 40% of infections.  SpO2 had FiO2 had even fewer data, covering fewer than 30% of infections.
```{r}
sepsis3 %>% 
  select(sofa_respiration, sofa_coagulation, sofa_liver, sofa_cardio, sofa_cns, sofa_renal) %>% 
  sapply(function(x) percent(sum(is.na(x))/length(x))) %>% 
  kable(caption = "Percent of missing data for each system", col.names = "% missing", full_width = F) %>% 
  kable_styling(bootstrap_options = "condensed", full_width = F)
sepsis3 %>% 
  count(data_type) %>% 
  mutate(pct = percent(n/nrow(sepsis3), accuracy = 0.01)) %>% 
  kable(caption = "Data composition of all suspected infections") %>% 
  kable_styling(bootstrap_options = "condensed", full_width = F)
```

Missing in SOFA < 2
```{r}
sepsis3 %>% 
  filter(sofa < 2) %>% 
  select(sofa_respiration, sofa_coagulation, sofa_liver, sofa_cardio, sofa_cns, sofa_renal) %>% 
  sapply(function(x) percent(sum(is.na(x))/length(x))) %>% 
  kable(caption = "Percent of missing data for each system", col.names = "% missing", full_width = F) %>% 
  kable_styling(bootstrap_options = "condensed", full_width = F)
sepsis3 %>% 
  filter(sofa < 2) %>% 
  count(data_type) %>% 
  mutate(pct = percent(n/sum(sepsis3$sofa < 2, na.rm = T), accuracy = 0.1)) %>% 
  kable(caption = "Data composition for SOFA < 2") %>% 
  kable_styling(bootstrap_options = "condensed", full_width = F)
```


# Rhee definition

1. For each CAM-ICU encounter, identify presumed serious infection .
    * Blood culture (BC) obtained _within $\pm1$ day of hopital admission date_, AND
    * $\ge 4$ QADs starting within $\pm2$ day of BC day.
        * __First QAD should be no later than discharge date, but no such limit for subsequent QADs__.
        * First QAD is the first day that patient receives a new antibiotic, which is not given in the prior 2 calendar days.
            * PO and IV/IM formulations of the same antibiotic are considered the same antibiotic EXCEPT for vacomycin. 
        * Subsequent QADs can be the same antibiotic, or different anitibiotics as long as the first dose of each antibiotic in the sequence is new.
            * A gap of a single calendar day between the same antibiotic count as QADs.
        * If the paitent died before 4 QADs, the paitent can still qualify for presumed infection as long as they have consecutive QADs until 1 day prior to death. 
        * There must be at least one new IV/IM antibiotic within the $\pm2$ day window of BC day.
2. For each presumed infection, if any of the acute organ dysfunction criteria was met within the $\pm2$ day window of BC day, it is a sepsis.
    * Vasopressor initiation (norepinephrine, dopamine, epinephrine, phenylephrine, or vasopressin)
        * note only IV were included following Rhee paper (page 23 in supplementary)
    * Initiation of mechanical ventilation
    * Doubling in creatinine level 
        * __We don't have eGFR data or ICD-9-CM code for end-stage kidney disease__.
    * Bilirubin level $\ge2.0$ mg/dL and doubling from baseline
    * Platelet count < 100x$10^3/\mu L$ and $\ge50\%$ decline from baseline (baseline must be < 100x$10^3/\mu L$) 
    * Lactate $\ge2.0$ mmol/L
    * Note
        * vasopressor and ventilation initiation if there were no vasopressors or mechanical ventilation admnistered on the prior calendar day
        * Since we limited BC day within $\pm1$ day of hopital admission date, all of our infections are "present-on-admission infection".  Baseline lab values were the best value during hospitalization, from 3 days before admission date to discharge date.

# Lab value range
* Creatinine
    * Some really big creatinine values > 20 mg/dL
    * 836 presumed infection had small baseline creatinine, even it's doubled it's still < 1.2, but this will satisfy Rhee definition of sepsis.
* A couple of bilirubin values > 50 mg/dL
* 17 presumed infection had baseline platelet count > 1000x$10^3/\mu L$
* FiO2 value < 21%
    * only 4% of the FiO2 < 21% had a matching Nasal O2 value.
    * If we find a way to utilize Nasal O2 data, it can add respiration data to ~900 ICU encouters. 
    * About 40 suspected infections had maximum FiO2 value < 21.

# Compare two definitions
Out of `r comma(nrow(cam_visits))` ICU encounters with CAM-ICU assessments, 

* Using __Sepsis-3__ definition
    * `r paste0(comma(nrow(sepsis3)), " (", percent(nrow(sepsis3)/nrow(cam_visits)), ")")` had suspected infections
    * __`r paste0(comma(sum(sepsis3$sepsis3 == 1)), " (", percent(sum(sepsis3$sepsis3 == 1)/nrow(sepsis3)), ")")`__ of suspected infections had a SOFA score $\ge 2$.  
    * Overall, __`r percent(sum(sepsis3$sepsis3 == 1)/nrow(cam_visits))`__ ICU encounters had sepsis, that is __`r comma(sepsis3 %>% filter(sepsis3 == 1) %>% distinct(grid) %>% nrow())`__ distinct patients.  
* Using __Rhee__ definition
    * `r paste0(comma(nrow(rhee_sepsis)), " (", percent(nrow(rhee_sepsis)/nrow(cam_visits)), ")")` had presumed infections
    * __`r paste0(comma(sum(rhee_sepsis$rhee == 1)), " (", percent(sum(rhee_sepsis$rhee == 1)/nrow(rhee_sepsis)), ")")`__ of presumed infections had acute organ dysfunction.  
    * Overall, __`r percent(sum(rhee_sepsis$rhee == 1)/nrow(cam_visits))`__ ICU encounters had sepsis, that is __`r comma(rhee_sepsis %>% filter(rhee == 1) %>% distinct(grid) %>% nrow())`__ distinct patients.

```{r}
sensitivity <- sepsis_comp %>% filter(rhee == 1, sepsis3 == 1) %>% nrow() / 
  sepsis_comp %>% filter(sepsis3 == 1) %>% nrow()
specificity <- sepsis_comp %>% filter(rhee == 0, sepsis3 == 0) %>% nrow() / 
  sepsis_comp %>% filter(sepsis3 == 0) %>% nrow()
```

## Sepsis-3 vs. Rhee 
Sensitivity and specificity of Rhee definition are `r percent(sensitivity)` and `r percent(specificity)` respectively.

```{r}
sepsis_comp %<>% 
  mutate(sepsis3 = factor(sepsis3, levels = c(1, 0)),
         rhee = factor(rhee, levels = c(1, 0)),
         sepsis3_infection = factor(sepsis3_infection, levels = c(1, 0)),
         rhee_infection = factor(rhee_infection, levels = c(1, 0))
         )
xtabs(~ rhee + sepsis3, data = sepsis_comp, addNA = T)
grid.newpage()
venn.plot <- draw.pairwise.venn(area1      = sepsis_comp %>% filter(sepsis3 == 1) %>% nrow(),
                                area2      = sepsis_comp %>% filter(rhee == 1) %>% nrow(),
                                cross.area = sepsis_comp %>% filter(rhee == 1, sepsis3 == 1) %>% nrow(),
                                category   = c("Sepsis-3", "Rhee"),
                                fill = c("blue", "red"),
                                lty = "blank",
                                ext.pos = 180)
```   

## Sepsis-3 infection vs. Rhee infection
```{r}
sensitivity <- sepsis_comp %>% filter(rhee_infection == 1, sepsis3_infection == 1) %>% nrow() / 
  sepsis_comp %>% filter(sepsis3_infection == 1) %>% nrow()
specificity <- sepsis_comp %>% filter(rhee_infection == 0, sepsis3_infection == 0) %>% nrow() / 
  sepsis_comp %>% filter(sepsis3_infection == 0) %>% nrow()
```


Sensitivity and specificity of Rhee definition of infection are `r percent(sensitivity)` and `r percent(specificity)` respectively.  In other words, all the infections identified by Rhee's definition are also infections under Sepsis-3 definition.
```{r}
xtabs(~ rhee_infection + sepsis3_infection, data = sepsis_comp, addNA = T)
grid.newpage()
venn.plot <- draw.pairwise.venn(area1      = sepsis_comp %>% filter(sepsis3_infection == 1) %>% nrow(),
                                area2      = sepsis_comp %>% filter(rhee_infection == 1) %>% nrow(),
                                cross.area = sepsis_comp %>% filter(rhee_infection == 1, sepsis3_infection == 1) %>% nrow(),
                                category   = c("Sepsis-3 Infection", "Rhee Infection"),
                                fill = c("blue", "red"),
                                lty = "blank",
                                cat.pos = c(0, 0),
                                cat.dist= c(-0.02, -0.02)
                                )

```

## Sepsis-3 vs. Rhee in infections identified by both definitions
```{r}
sensitivity <- sepsis_comp %>% filter(rhee_infection == 1, rhee == 1, sepsis3 == 1) %>% nrow() / 
  sepsis_comp %>% filter(rhee_infection == 1, sepsis3 == 1) %>% nrow()
specificity <- sepsis_comp %>% filter(rhee_infection == 1, rhee == 0, sepsis3 == 0) %>% nrow() / 
  sepsis_comp %>% filter(rhee_infection == 1, sepsis3 == 0) %>% nrow()
```
In these `r comma(sepsis_comp %>% filter(rhee_infection == 1) %>% nrow())` infections, Sensitivity and specificity of Rhee definition are `r percent(sensitivity)` and `r percent(specificity)` respectively.  
So even when we only consider infections identified by both definitions, Sepsis-3 still identify more sepsis compared to Rhee. 
```{r}
sepsis_comp %>% 
  filter(rhee_infection == 1) %>% 
  with(xtabs(~rhee+sepsis3))
grid.newpage()
venn.plot <- draw.pairwise.venn(area1      = sepsis_comp %>% filter(rhee_infection == 1, sepsis3 == 1) %>% nrow(),
                                area2      = sepsis_comp %>% filter(rhee_infection == 1, rhee == 1) %>% nrow(),
                                cross.area = sepsis_comp %>% filter(rhee_infection == 1, sepsis3 == 1, rhee == 1) %>% nrow(),
                                category   = c("Sepsis-3", "Rhee"),
                                fill = c("blue", "red"),
                                lty = "blank",
                                ext.pos = 180
)

```

# Patient Location
The blood gas lab data has location data associated with it.  70% of CAM-ICU encounters had some location data.  Maybe we can use the location data to help to identify ICU stays.  
Below is location counts by hospital days.
```{r}
loc_ct <- read_csv("../output/patient_cam_visit_location_count.csv")
loc_ct
```

