---
title: "BioVU Missing CAM"
author: "Fei Ding"
date: "April 3, 2019"
output:
  html_document:
    toc: true
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
library(tidyverse)
library(readxl)
library(Hmisc)
library(magrittr)
library(scales)
library(lubridate)
library(kableExtra)

discharge_raw <- read_csv("../data/Samuels_Delirium_DISCHARGE_20180718.csv")
daily_status <- read_csv("../output/daily_status_20190319.csv")
hosp2 <- read_excel("../data/Hospitalization_dates_DCS_V01.xlsx", 
                    sheet = "Hospitalization_dates_DCS_V01",
                    na = c("", "NA"))
names(discharge_raw) <- tolower(names(discharge_raw))
discharge_raw1 <- discharge_raw %>% 
  mutate(adm_dttm = mdy_hms(admiss_date),
         adm_date = as_date(adm_dttm),
         dc_dttm = mdy_hms(discharge_date),
         dc_date = as_date(dc_dttm)) 

discharge_raw1  %<>% 
  filter(year(adm_date) >= 2004) %>%  # CAM data spans from 2005 to 2017
  distinct(grid, adm_date, dc_date) %>% 
  arrange(grid, adm_date, dc_date) %>% 
  group_by(grid) %>% 
  mutate(next_adm = lead(adm_date), 
         next_dc = lead(dc_date),
         pre_adm = lag(adm_date),
         pre_dc = lag(dc_date)
  ) %>% 
  ungroup() 
discharge_raw2  <- discharge_raw1 %>% 
  filter(!is.na(dc_date)) %>% 
  group_by(grid) %>% 
  mutate(next_adm = lead(adm_date), 
         next_dc = lead(dc_date),
         pre_adm = lag(adm_date),
         pre_dc = lag(dc_date)
  ) %>% 
  ungroup()
discharge_raw2 %<>% 
  filter(adm_date != pre_dc | is.na(pre_dc)) %>% 
  mutate(
    dc_date = case_when(
      dc_date == next_adm ~ next_dc,
      T ~ dc_date)
  )
discharge_raw2 %<>% 
  group_by(grid) %>% 
  mutate(adm_id = 1:n(),
         adm_n = n()) %>% 
  ungroup() %>% 
  select(grid, adm_n, adm_id, adm_date, dc_date)
daily_status_f1 <- sqldf::sqldf('SELECT *
        FROM daily_status as t1
        LEFT JOIN discharge_raw2 as t2 ON t1.grid = t2.grid AND dt BETWEEN adm_date AND dc_date
        ')

names(daily_status_f1)[12] <- "grid.1"
daily_status_f1 <- as_tibble(daily_status_f1) %>% 
  arrange(grid, dt) %>% 
  mutate(
    adm_day = dt - adm_date + 1
  )



hosp2 %<>% 
  mutate(
    adm_date = as_date(adm_date),
    dc_date = as_date(dc_date),
    dc_date1 = case_when(
      !grepl("drop", `revised dc_date`) ~
        ymd(19000101) + as.numeric(`revised dc_date`) - 2)
  )
hosp3 <- hosp2 %>%  
  filter(!is.na(dc_date1)) %>% 
  distinct(grid, adm_date, dc_date1) %>% 
  rename(dc_date = dc_date1) %>% 
  arrange(grid, adm_date, dc_date) %>% 
  group_by(grid) %>% 
  mutate(next_adm = lead(adm_date), 
         next_dc = lead(dc_date),
         pre_adm = lag(adm_date),
         pre_dc = lag(dc_date)) %>% 
  ungroup()
hosp3 %<>%  
  filter(adm_date > pre_adm | is.na(pre_adm)) %>% 
  filter(adm_date > pre_dc | is.na(pre_dc)) %>% 
  mutate(
    dc_date = case_when(
      adm_date == next_adm ~ next_dc,
      dc_date == next_adm ~ next_dc,
      T ~ dc_date)
  ) %>% 
  arrange(grid, adm_date, dc_date) %>% 
  group_by(grid) %>% 
  mutate(next_adm = lead(adm_date), 
         next_dc = lead(dc_date),
         pre_adm = lag(adm_date),
         pre_dc = lag(dc_date)) %>% 
  ungroup() 
hosp4 <- hosp3 %>%
  select(grid:dc_date) %>% 
  group_by(grid) %>% 
  mutate(adm_id = 1:n(),
         adm_n = n()) %>% 
  ungroup()
daily_status_f2 <- sqldf::sqldf('SELECT *
        FROM daily_status as t1
        LEFT JOIN hosp4 as t2 ON t1.grid = t2.grid AND dt BETWEEN adm_date AND dc_date
        ') #%>% 
names(daily_status_f2)[12] <- "grid.1"
daily_status_f2 <- as_tibble(daily_status_f2) %>% 
  arrange(grid, dt) %>% 
  mutate(
    adm_day = dt - adm_date + 1
  )
```

## Data processing

There are two different hospital admission/discharge files.  The old one *Samuels_Delirium_DISCHARGE_20180718.csv* had more records and patients, while the new one *Hospitalization_dates_DCS_V01.xlsx* had fewer.

They were processed in same way: 

* Missing discharge records were removed.
* Overlapped admissions were combined.
* For the old file, admissions starting from 2004 were kept

The old file has almost twice of admission/discharge records compared to the new file, and `r n_distinct(discharge_raw2$grid) - n_distinct(hosp4$grid)` more patients.

```{r}
discharge_raw2 %>% 
  count(year(adm_date)) %>% 
  full_join(
    hosp4 %>% 
      count(year(adm_date)),
    by = "year(adm_date)",
    suffix = c(".old", ".new")
  ) %>% 
  add_row(`year(adm_date)` = "Total", n.old = nrow(discharge_raw2), n.new = nrow(hosp4)) %>% 
  kable(caption = "Years of Admission")  %>% 
  kable_styling(bootstrap_options = c("condensed", "bordered"), 
                full_width = F,
                position = "left") %>% 
  column_spec(1, bold = T) %>% 
  row_spec(15, bold = T)

  
```

As a reminder, last time we got daily status as following.  
```{r}
daily_status %>% 
  count(status.today) %>% 
  add_row(status.today = "Total", nn = nrow(daily_status)) %>% 
  kable(caption = "Daily Status") %>% 
  kable_styling(bootstrap_options = c("condensed", "bordered"), 
                full_width = F,
                position = "left") %>% 
  column_spec(1, bold = T) %>% 
  row_spec(6, bold = T)
```


After merging daily status and admission/discharge records, below are the assessment days without a matching admission/discharge record.
```{r}
daily_status_f1 %>% 
  filter(is.na(adm_date)) %>% # 108,998
  count(status.today) %>% 
  full_join(
    daily_status_f2 %>% 
  filter(is.na(adm_date)) %>% # 60,114 days
  count(status.today),
  by = "status.today", 
  suffix = c(".old", ".new")
  ) %>% 
  add_case(status.today = "Total", 
           nn.old = sum(is.na(daily_status_f1$adm_date)), 
           nn.new = sum(is.na(daily_status_f2$adm_date))
                        ) %>% 
  kable(caption = "Assessment Days without matching admission") %>% 
    kable_styling(bootstrap_options = c("condensed", "bordered"), 
                full_width = F,
                position = "left") %>% 
  column_spec(1, bold = T) %>% 
  row_spec(6, bold = T)
```

## Missing CAM vs. hospital admission day
Only assessment days with a matching admission/discharge record was plotted here.

For the bar plot, admission day was capped at day 50.
```{r}
## make plots
daily_status_f1 %>% 
  filter(adm_day <= 50) %>% 
  ggplot() +
  geom_bar(aes(x = adm_day, fill = status.today)) +
  labs(title = "Missing CAM vs. Hospital admission day", x = "Hospital admission day", y = "count of visits",
       subtitle = "Using old admission/discharge data")

daily_status_f1 %>% 
  count(adm_day) %>% 
  right_join(
    daily_status_f1 %>% 
      filter(status.today == "Unknown: RASS only") %>% 
      count(adm_day),
    by = "adm_day"
  ) %>% 
  mutate(pct = nn.y/nn.x) %>% 
  ggplot(aes(x = adm_day, y = pct)) +
  geom_point() +
  geom_line() +
  labs(title = "Missing CAM % vs. Hospital admission day", y = "Percent of days Missing CAM", x = "Hospital admission day",
       subtitle = "Using old admission/discharge data")


daily_status_f2 %>% 
  filter(adm_day <= 50) %>% 
  ggplot() +
  geom_bar(aes(x = adm_day, fill = status.today)) +
    labs(title = "Missing CAM vs. Hospital admission day", x = "Hospital admission day", y = "count of visits",
       subtitle = "Using old admission/discharge data")

  
daily_status_f2 %>% 
  count(adm_day) %>% 
  right_join(
    daily_status_f2 %>% 
      filter(status.today == "Unknown: RASS only") %>% 
      count(adm_day),
    by = "adm_day"
  ) %>% 
  mutate(pct = nn.y/nn.x) %>% 
  ggplot(aes(x = adm_day, y = pct)) +
  geom_point() +
  geom_line() +
 labs(title = "Missing CAM % vs. Hospital admission day", y = "Percent of days Missing CAM", x = "Hospital admission day",
       subtitle = "Using old admission/discharge data")


```

## Where do RASS-only days come from?
Are they from hospital visits with no CAM-ICU at all or are they from hospital visits with some CAM?
Only assessment days with a matching admission/discharge record was tabulated here.

```{r}
c1 <- daily_status_f1 %>% 
  filter(status.today == "Unknown: RASS only", !is.na(adm_id)) %>% 
  left_join(   
    daily_status_f1 %>% 
      filter(!is.na(adm_id)) %>%
      mutate(n_cam = n_cam_neg + n_cam_pos) %>% 
      group_by(grid, adm_id) %>% 
      summarise(some_cam = max(n_cam) > 0),
    by = c("grid", "adm_id")
  ) %>% 
  count(some_cam) %>%
  add_case(some_cam = "Total", 
           nn = sum(daily_status_f1$status.today == "Unknown: RASS only" & !is.na(daily_status_f1$adm_id)))
  

c2 <- daily_status_f2 %>% 
  filter(status.today == "Unknown: RASS only", !is.na(adm_id)) %>% 
  left_join(   
    daily_status_f2 %>% 
      filter(!is.na(adm_id)) %>%
      mutate(n_cam = n_cam_neg + n_cam_pos) %>% 
      group_by(grid, adm_id) %>% 
      summarise(some_cam = max(n_cam) > 0),
    by = c("grid", "adm_id")
  ) %>% 
  count(some_cam)%>%
  add_case(some_cam = "Total", 
           nn = sum(daily_status_f2$status.today == "Unknown: RASS only" & !is.na(daily_status_f2$adm_id)))
  
c1 %>% 
  full_join(c2,
            by = "some_cam",
            suffix = c(".old", ".new")) %>% 
  kable(caption = "RASS only days") %>% 
    kable_styling(bootstrap_options = c("condensed", "bordered"), 
                full_width = F,
                position = "left") %>% 
  column_spec(1, bold = T) %>% 
  row_spec(3, bold = T)
```
## Example of RASS only days
```{r}
tb1 <- daily_status_f1 %>% 
  select(grid, adm_n, adm_id, adm_date, dc_date, dt, adm_day,status.today) %>% 
  filter(grid %in% c("R200000649", "R200006516")) #%>% 
  #   mutate(
  #   status.today = cell_spec(
  #     status.today, "html",
  #     color = "white",
  #     background  = case_when(
  #       status.today == "Delirious" ~ "#D55E00", 
  #       status.today == "Normal" ~ "#56B4E9",
  #       status.today == "Comatose" ~ "black",
  #       status.today == "Unknown: RASS only" ~ "gray")
  #     )
  #   ) %>% 
  #  kable(format = "html", escape = F, caption = "Old admission data") %>% 
  # kable_styling(bootstrap_options = c("condensed", "bordered"), 
  #               full_width = F,
  #               position = "left") 

tb2 <- daily_status_f2 %>% 
  select(grid, adm_n, adm_id, adm_date, dc_date, dt, adm_day,status.today) %>% 
  filter(grid %in% c("R200000649", "R200006516")) #%>% 
  #   mutate(
  #   status.today = cell_spec(
  #     status.today, "html",
  #     color = "white",
  #     background  = case_when(
  #       status.today == "Delirious" ~ "#D55E00", 
  #       status.today == "Normal" ~ "#56B4E9",
  #       status.today == "Comatose" ~ "black",
  #       status.today == "Unknown: RASS only" ~ "gray")
  #     )
  #   ) %>% 
  #  kable(format = "html", escape = F, caption = "New admission data") %>% 
  # kable_styling(bootstrap_options = c("condensed", "bordered"), 
  #               full_width = F,
  #               position = "left") 

tb <- tb1 %>% 
  full_join(tb2,
            by = c("grid", "dt", "status.today"),
            suffix = c(".old", ".new")
  ) %>% 
  select(grid, dt, status.today, ends_with(".old"), ends_with((".new"))) 
rowid1 <- with(tb,   which(is.na(adm_n.new) & !is.na(adm_n.old))) 
rowid2 <- with(tb,   which(is.na(adm_n.new) & is.na(adm_n.old))) 
tb %>% 
  select(grid:adm_id.new) %>% 
  mutate(
    status.today = cell_spec(
      status.today, "html",
      color = "white",
      background  = case_when(
        status.today == "Delirious" ~ "#D55E00", 
        status.today == "Normal" ~ "#56B4E9",
        status.today == "Comatose" ~ "black",
        status.today == "Unknown: RASS only" ~ "gray")
      )
    ) %>% 
   kable(format = "html", escape = F, caption = "New admission data") %>% 
  kable_styling(bootstrap_options = c("condensed", "bordered"), 
                full_width = F,
                position = "left") %>% 
  row_spec(rowid1, background = "lightyellow") %>% 
  row_spec(rowid2, background = "lightblue")
```

