---
title: "Identify Sepsis"
author: "Fei Ding"
date: "December 3, 2019"
output:
  html_document:
    df_print: paged
---




```{r setup, include = FALSE}
library(tidyverse)
library(scales)
library(kableExtra)
library(VennDiagram)
library(magrittr)

knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)
dob_dis <- read_csv("../output/dob_discrepancy.csv") 
#cam_visits <- read_csv("../output/cam_stay_20190925.csv") 
sepsis3 <- read_csv("../output/sepsis3_20191014.csv")
rhee_sepsis <- read_csv("../output/sepsis_rhee_20191203.csv")
sepsis_comp <- read_csv("../output/sepsis_compare_20191203.csv")

sepsis_comp %>% 
  semi_join(dob_dis)
sepsis_comp %>% 
  semi_join(dob_dis) %>% 
  distinct(grid)

```

# Update from last report
1. Add eGFR and end-stage kidney disease code to Rhee definition
2. Compare Rhee and Sepsis-3 definition to Sepsis code.


_Note that 15 patients with DOB discrepancy > 10 days were removed from our final results._
```{r}
#cam_visits %<>% anti_join(dob_dis, by = "grid")
sepsis3 %<>% anti_join(dob_dis, by = "grid")
rhee_sepsis %<>% anti_join(dob_dis, by = "grid")
sepsis_comp %<>% anti_join(dob_dis, by = "grid")

```



# Rhee definition

1. For each CAM-ICU encounter, identify presumed serious infection .
    * Blood culture (BC) obtained _within $\pm1$ day of hopital admission date_, AND
    * $\ge 4$ QADs starting within $\pm2$ day of BC day.
        * __First QAD should be no later than discharge date, but no such limit for subsequent QADs__.
        * First QAD is the first day that patient receives a new antibiotic, which is not given in the prior 2 calendar days.
            * PO and IV/IM formulations of the same antibiotic are considered the same antibiotic EXCEPT for vacomycin. 
        * Subsequent QADs can be the same antibiotic, or different anitibiotics as long as the first dose of each antibiotic in the sequence is new.
            * A gap of a single calendar day between the same antibiotic count as QADs.
        * If the paitent died before 4 QADs, the paitent can still qualify for presumed infection as long as they have consecutive QADs until 1 day prior to death. 
        * There must be at least one new IV/IM antibiotic within the $\pm2$ day window of BC day.
2. For each presumed infection, if any of the acute organ dysfunction criteria was met within the $\pm2$ day window of BC day, it is a sepsis.
    * Vasopressor initiation (norepinephrine, dopamine, epinephrine, phenylephrine, or vasopressin)
        * note only IV were included following Rhee paper (page 23 in supplementary)
    * Initiation of mechanical ventilation
    * Doubling in creatinine level or decrease by $\ge50\%$ decline eGFR relative to baseline (excluding patients with ICD code for end-stage kidney disease).
        * eGFR was calcualted based on CKD-EPI equation in Table 1 on this [webpage](https://www.niddk.nih.gov/health-information/communication-programs/nkdep/laboratory-evaluation/glomerular-filtration-rate/estimating)
        * __An admission was considered end-stage kidney disease thus excluded only if there is an ESRD code during that admission (for consistency, counted from 3 days before admission date to discharge date).__
            * About 123 patients had admission satisfying this condition but who had ESRD code for earlier admission.
        * Race were dichotomized as Black vs. White and others. 
            * __B (Black). When a GRID has conflicting races, if any of the recorded race include "B", this patient is also considered Black.__
            * __White and other include A (Asian),  W (White), I (Islander? American Indian?), O(Other?) ,D(?), N(Native American?)__
            * U is treated as missing.
        * Note that eGFR > 60 was set to 60 following Rhee's paper.
    * Bilirubin level $\ge2.0$ mg/dL and doubling from baseline
    * Platelet count < 100x$10^3/\mu L$ and $\ge50\%$ decline from baseline (baseline must be < 100x$10^3/\mu L$) 
    * Lactate $\ge2.0$ mmol/L
    * Note
        * vasopressor and ventilation initiation if there were no vasopressors or mechanical ventilation admnistered on the prior calendar day
        * Since we limited BC day within $\pm1$ day of hopital admission date, all of our infections are "present-on-admission infection".  Baseline lab values were the best value during hospitalization, from 3 days before admission date to discharge date.


# Sepsis-3 definition

**NOTE:** Deviation from original definition were **bold (need more discussion)** or _italic (decided deviation)_.

1. For each CAM-ICU encounter, identify the first episode of suspected infection _within $\pm1$ day of hopital admission date_.
    * If the antibiotic was given first, the culture sampling must have been obtained within 1 day.
    * If the culture sampling was first, the antibiotic must have been ordered within 3 days.
    * The onset date of infection was the date on which the first of these two events occured.
    * __We used only blood culutre instead of body fluid cultures (blood, urine, cerebrospinal fluid, etc)__.
2. For each suspected infection, find worst daily SOFA score for the time window from 2 day before to 1 day after the onset date of infection.
    * Central nervous system SOFA score calculated _based on RASS Method C in the Modified SOFA score paper_.  
    * Respiration SOFA score calculated using both PaO~2~/FiO~2~(PF) ratio _and SpO~2~/FiO~2~(SF) ratio_.
        * FiO~2~ values > 100 were excluded. 1 < FiO~2~ <= 100 were converted to 0-1. 
        * SpO~2~ values > 100 were converted to 100.
    * __We used only creatinine data for renal SOFA, no urine output data__.
    * __For vasopressor, there is a footnote "Catecholamine doses are given as $\mu g/kg/min$ for at least 1 hour", do we need to limit vasopressor to IV only (except dobutamine)?__
    * Note
        * All the lab/drug dates are from 3 days before admission date to discharge date. 
        * All non-numerical lab values were converted to upper/lower limit or mean of the range.  
   
    
|SOFA score|PF or SF ratio and Ventilation | Platelets, x$10^3/\mu L$| Bilirubin, mg/dL| Cardiovascular| Rass score | Creatinine, mg/dL |
--|-------------------------------------|----------|-----------|----------------------------------|--------------|---|----|
0 | PF $\ge$ 400 or SF $\ge$ 512          | $\ge$150 | <1.2     | MAP $\ge$70 mmHg                  |$\ge$ 0   | <1.2    |  
1 | PF < 400 or SF < 512                  | <150     | 1.2-1.9  | MAP < 70 mmHg                     | -1       | 1.2-1.9 |
2 | PF < 300 or SF < 357                  | <100     | 2.0-5.9  | any dopamine or dubutamine        | -2       | 2.0-3.4 |
3 | PF < 200 or SF < 214 with ventilation | <50      | 6.0-11.9 | any epinephrine or norepinephrine | -3       | 3.5-4.9 |
4 | PF < 100 or SF < 89 with ventilation  | <20      | >12.0    |                                   | -4 to -5 | >5.0    |

3. Suspected infection with a SOFA score $\ge2$ is considered sepsis __assuming baseline SOFA score is zero__.




# Compare two definitions with sepsis code
Out of `r comma(nrow(sepsis_comp))` ICU encounters with CAM-ICU assessments, 

* Using __Rhee__ definition
    * `r paste0(comma(nrow(rhee_sepsis)), " (", percent(nrow(rhee_sepsis)/nrow(sepsis_comp)), ")")` had presumed infections
    * __`r paste0(comma(sum(rhee_sepsis$rhee == 1)), " (", percent(sum(rhee_sepsis$rhee == 1)/nrow(rhee_sepsis)), ")")`__ of presumed infections had acute organ dysfunction.  It's fewer compared to last report because we excluded patients with end-stage kidney disease.
    * Overall, __`r percent(sum(rhee_sepsis$rhee == 1)/nrow(sepsis_comp))`__ ICU encounters had sepsis, that is __`r comma(rhee_sepsis %>% filter(rhee == 1) %>% distinct(grid) %>% nrow())`__ distinct patients.
* Using __Sepsis-3__ definition
    * `r paste0(comma(nrow(sepsis3)), " (", percent(nrow(sepsis3)/nrow(sepsis_comp)), ")")` had suspected infections
    * __`r paste0(comma(sum(sepsis3$sepsis3 == 1)), " (", percent(sum(sepsis3$sepsis3 == 1)/nrow(sepsis3)), ")")`__ of suspected infections had a SOFA score $\ge 2$.  
    * Overall, __`r percent(sum(sepsis3$sepsis3 == 1)/nrow(sepsis_comp))`__ ICU encounters had sepsis, that is __`r comma(sepsis3 %>% filter(sepsis3 == 1) %>% distinct(grid) %>% nrow())`__ distinct patients.  
* Using __Sepsis code__
    * __`r comma(sum(sepsis_comp$sepsis_code_adm == 1))`__ ICU encounters had sepsis code during hospitalization (from 1 day before admission date to discharge date), that is __`r comma(sepsis_comp %>% filter(sepsis_code_adm == 1) %>% distinct(grid) %>% nrow())`__ patients. 
    * __`r comma(sum(sepsis_comp$sepsis_code_w1d == 1))`__ ICU encounters had sepsis code within $\pm1$ day of hopital admission date, that is __`r comma(sepsis_comp %>% filter(sepsis_code_w1d == 1) %>% distinct(grid) %>% nrow())`__ patients
* Using either __Rhee or Sepsis-3__ definition
    *  __`r paste0(comma(sepsis_comp %>% filter(sepsis3 == 1 | rhee == 1) %>% nrow()), " (", percent(sepsis_comp %>% filter(sepsis3 == 1 | rhee == 1) %>% nrow()/nrow(sepsis_comp)), ")")`__ ICU encounters had sepsis, that is __`r comma(sepsis_comp %>% filter(sepsis3 == 1 | rhee == 1) %>% distinct(grid) %>% nrow())`__ distinct patients. 
* Using __Rhee or Sepsis-3 definition or sepsis code__ within $\pm1$ day of hopital admission
    *  __`r paste0(comma(sepsis_comp %>% filter(sepsis3 == 1 | rhee == 1 | sepsis_code_w1d == 1) %>% nrow()), " (", percent(sepsis_comp %>% filter(sepsis3 == 1 | rhee == 1| sepsis_code_w1d == 1) %>% nrow()/nrow(sepsis_comp)), ")")`__ ICU encounters had sepsis, that is __`r comma(sepsis_comp %>% filter(sepsis3 == 1 | rhee == 1| sepsis_code_w1d == 1) %>% distinct(grid) %>% nrow())`__ distinct patients. 
* Lists of GRIDs with sepsis were provided in _Output/sepsis_grids_20191204.xlsx_ for Rhee definition only, Rhee or Sepsis-3 definition, Rhee or Sepsis-3 or Sepsis code.

```{r}
sepsis_comp %<>% 
  mutate(sepsis3 = factor(sepsis3, levels = c(1, 0)),
         rhee = factor(rhee, levels = c(1, 0)),
         sepsis_code_adm = factor(sepsis_code_adm, levels = c(1, 0)),
         sepsis_code_w1d = factor(sepsis_code_w1d, levels = c(1, 0))
         )
sensitivity <- sepsis_comp %>% filter(rhee == 1, sepsis3 == 1) %>% nrow() / 
  sepsis_comp %>% filter(sepsis3 == 1) %>% nrow()
specificity <- sepsis_comp %>% filter(rhee == 0, sepsis3 == 0) %>% nrow() / 
  sepsis_comp %>% filter(sepsis3 == 0) %>% nrow()
```

## Sepsis-3 vs. Rhee 
Sensitivity and specificity of Rhee definition are `r percent(sensitivity)` and `r percent(specificity)` respectively.

```{r}
sepsis_comp %<>% 
  mutate(sepsis3 = factor(sepsis3, levels = c(1, 0)),
         rhee = factor(rhee, levels = c(1, 0)),
         sepsis3_infection = factor(sepsis3_infection, levels = c(1, 0)),
         rhee_infection = factor(rhee_infection, levels = c(1, 0))
         )
xtabs(~ rhee + sepsis3, data = sepsis_comp, addNA = T)
grid.newpage()
venn.plot <- draw.pairwise.venn(area1      = sepsis_comp %>% filter(sepsis3 == 1) %>% nrow(),
                                area2      = sepsis_comp %>% filter(rhee == 1) %>% nrow(),
                                cross.area = sepsis_comp %>% filter(rhee == 1, sepsis3 == 1) %>% nrow(),
                                category   = c("Sepsis-3", "Rhee"),
                                fill = c("blue", "red"),
                                lty = "blank",
                                ext.pos = 180)
```   

## Rhee vs. Sepsis code 
### Sepsis code during hospitalization

```{r}
sensitivity <- sepsis_comp %>% filter(rhee == 1, sepsis_code_adm == 1) %>% nrow() / 
  sepsis_comp %>% filter(sepsis_code_adm == 1) %>% nrow()
specificity <- sepsis_comp %>% filter(rhee == 0, sepsis_code_adm == 0) %>% nrow() / 
  sepsis_comp %>% filter(sepsis_code_adm == 0) %>% nrow()
```


Sensitivity and specificity of Rhee definition are `r percent(sensitivity)` and `r percent(specificity)` respectively.  
```{r}
xtabs(~ rhee + sepsis_code_adm, data = sepsis_comp, addNA = T)
grid.newpage()
venn.plot <- draw.pairwise.venn(area1      = sepsis_comp %>% filter(sepsis_code_adm == 1) %>% nrow(),
                                area2      = sepsis_comp %>% filter(rhee == 1) %>% nrow(),
                                cross.area = sepsis_comp %>% filter(rhee == 1, sepsis_code_adm == 1) %>% nrow(),
                                category   = c("Sepsis code", "Rhee"),
                                fill = c("blue", "red"),
                                lty = "blank"
        
                                )

```

### Sepsis code within $\pm1$ day of hopital admission date

```{r}
sensitivity <- sepsis_comp %>% filter(rhee == 1, sepsis_code_w1d == 1) %>% nrow() / 
  sepsis_comp %>% filter(sepsis_code_w1d == 1) %>% nrow()
specificity <- sepsis_comp %>% filter(rhee == 0, sepsis_code_w1d == 0) %>% nrow() / 
  sepsis_comp %>% filter(sepsis_code_w1d == 0) %>% nrow()
```


Sensitivity and specificity of Rhee definition are `r percent(sensitivity)` and `r percent(specificity)` respectively.  
```{r}
xtabs(~ rhee + sepsis_code_w1d, data = sepsis_comp, addNA = T)
grid.newpage()
venn.plot <- draw.pairwise.venn(area1      = sepsis_comp %>% filter(sepsis_code_w1d == 1) %>% nrow(),
                                area2      = sepsis_comp %>% filter(rhee == 1) %>% nrow(),
                                cross.area = sepsis_comp %>% filter(rhee == 1, sepsis_code_w1d == 1) %>% nrow(),
                                category   = c("Sepsis code", "Rhee"),
                                fill = c("blue", "red"),
                                lty = "blank"
                                )

```

## Sepsis-3 vs. Sepsis code 
### Sepsis code during hospitalization

```{r}
sensitivity <- sepsis_comp %>% filter(sepsis3 == 1, sepsis_code_adm == 1) %>% nrow() / 
  sepsis_comp %>% filter(sepsis_code_adm == 1) %>% nrow()
specificity <- sepsis_comp %>% filter(sepsis3 == 0, sepsis_code_adm == 0) %>% nrow() / 
  sepsis_comp %>% filter(sepsis_code_adm == 0) %>% nrow()
```


Sensitivity and specificity of sepsis-3 definition are `r percent(sensitivity)` and `r percent(specificity)` respectively.  
```{r}
xtabs(~ sepsis3 + sepsis_code_adm, data = sepsis_comp, addNA = T)
grid.newpage()
venn.plot <- draw.pairwise.venn(area1      = sepsis_comp %>% filter(sepsis_code_adm == 1) %>% nrow(),
                                area2      = sepsis_comp %>% filter(sepsis3 == 1) %>% nrow(),
                                cross.area = sepsis_comp %>% filter(sepsis3 == 1, sepsis_code_adm == 1) %>% nrow(),
                                category   = c("Sepsis code", "Sepsis-3"),
                                fill = c("blue", "red"),
                                lty = "blank"
                                )

```

### Sepsis code within $\pm1$ day of hopital admission date

```{r}
sensitivity <- sepsis_comp %>% filter(sepsis3 == 1, sepsis_code_w1d == 1) %>% nrow() / 
  sepsis_comp %>% filter(sepsis_code_w1d == 1) %>% nrow()
specificity <- sepsis_comp %>% filter(sepsis3 == 0, sepsis_code_w1d == 0) %>% nrow() / 
  sepsis_comp %>% filter(sepsis_code_w1d == 0) %>% nrow()
```


Sensitivity and specificity of Sepsis-3 definition are `r percent(sensitivity)` and `r percent(specificity)` respectively.  
```{r}
xtabs(~ sepsis3 + sepsis_code_w1d, data = sepsis_comp, addNA = T)
grid.newpage()
venn.plot <- draw.pairwise.venn(area1      = sepsis_comp %>% filter(sepsis_code_w1d == 1) %>% nrow(),
                                area2      = sepsis_comp %>% filter(sepsis3 == 1) %>% nrow(),
                                cross.area = sepsis_comp %>% filter(sepsis3 == 1, sepsis_code_w1d == 1) %>% nrow(),
                                category   = c("Sepsis code", "Sepsis-3"),
                                fill = c("blue", "red"),
                                lty = "blank"
                                )

```

## Rhee vs. Sepsis-3 vs. Sepsis code within $\pm1$ day of hopital admission date
```{r}
grid.newpage()
venn.plot <- draw.triple.venn(area1      = sepsis_comp %>% filter(sepsis_code_w1d == 1) %>% nrow(),
                                area2      = sepsis_comp %>% filter(sepsis3 == 1) %>% nrow(),
                                area3      = sepsis_comp %>% filter(rhee == 1) %>% nrow(),
                              n12 = sepsis_comp %>% filter(sepsis3 == 1, sepsis_code_w1d == 1) %>% nrow(),
                              n23 = sepsis_comp %>% filter(rhee == 1, sepsis3 == 1) %>% nrow(),
                              n13 = sepsis_comp %>% filter(rhee == 1, sepsis_code_w1d == 1) %>% nrow(),
                                n123 = sepsis_comp %>% filter(rhee == 1, sepsis3 == 1, sepsis_code_w1d == 1) %>% nrow(),
                                category   = c("Sepsis code", "Sepsis-3", "Rhee"),
                                fill = c("green", "blue", "red"),
                                lty = "blank"
                                )

```

Of the 2998 ICU encounters with Sepsis code but not identified as sepsis by either Sepsis-3 or Rhee definition, 2659 were not identified as infection by either definition. The other 339 had sepsis-3 definition, but SOFA < 2 due to missing data.  Of the 339 encounters, 99 also had Rhee definition, but they did not meet acute organ dysfunction criteria.
```{r}
## check sepsis-code positive other negative -------------
sepsis_comp %>% 
  filter(sepsis_code_w1d == 1, sepsis3 != 1, rhee != 1) %>% 
  count(sepsis3_infection, rhee_infection) %>% 
  kable() %>% 
  kable_styling("condensed", full_width = F)
#' 240 had both rhee and sepsis-3 infecgtion, 99 had sepsis-3 infection only
sepsis_comp %>% 
  filter(sepsis_code_w1d == 1, sepsis3 != 1, rhee != 1, sepsis3_infection == 1) %>% 
  count(data_type) %>% 
  kable() %>% 
  kable_styling("condensed", full_width = F)
```


# Lab value range
* Creatinine
    * Some really big creatinine values > 20 mg/dL
    * 825 presumed infection had small baseline creatinine, even it's doubled it's still < 1.2, but this will satisfy Rhee definition of sepsis.
* A couple of bilirubin values > 50 mg/dL
* 17 presumed infection had baseline platelet count > 1000x$10^3/\mu L$
* FiO2 value < 21%
    * only 4% of the FiO2 < 21% had a matching Nasal O2 value.
    * If we find a way to utilize Nasal O2 data, it can add respiration data to ~900 ICU encouters. 
    * About 40 suspected infections had maximum FiO2 value < 21.
