---
title: "BioVU CAM-ICU visit summary"
author: "Fei Ding"
date: "July 16, 2019"
output:
  html_document:
    df_print: paged
    
---

## Visit summary
```{r setup, include = FALSE}
library(tidyverse)
library(readxl)
library(lubridate)
library(scales)
library(kableExtra)

knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)

cam_rass_daily <- read_csv("../output/daily_status_20190617.csv")
cam_gap <- read_csv("../output/cam_gap_20190618.csv")
daily_status_f1 <- sqldf::sqldf('SELECT *
        FROM cam_rass_daily as t1
        INNER JOIN cam_gap as t2 
        ON t1.grid = t2.grid AND dt BETWEEN first_cam AND last_cam') %>% 
  as_tibble()
visit_status1 <- daily_status_f1 %>% 
  filter(status.today %in% c("Comatose", "Delirious", "Normal", "Unkown: conflicting CAM")) %>% 
  group_by(grid, adm_id) %>% 
  count(status.today) %>% 
  ungroup() %>% 
  spread(key = status.today, value = n, fill = 0L) 

visit_status2 <- visit_status1 %>% 
  full_join(cam_gap, by = c("grid", "adm_id")) %>% 
  mutate(any_coma = ifelse(Comatose > 0, 1, 0),
         any_del = ifelse(Delirious > 0, 1, 0),
         all_norm = ifelse(Normal == bt_days, 1, 0),
         norm_un = ifelse(any_coma + any_del + all_norm == 0, 1, 0)
         )
coma_q <- visit_status2 %>% 
  filter(any_coma > 0) %>% 
  pull(Comatose) %>% 
  quantile()
del_q <- visit_status2 %>% 
  filter(any_del > 0) %>% 
  pull(Delirious) %>% 
  quantile()
```

As stated in the last report, if we ignore CAM-ICUs without a matching admission/discharge record, there are `r comma(nrow(cam_gap))` CAM-ICU visits, and `r paste0(comma(sum(cam_gap$n_hole == 0)), " (", percent(sum(cam_gap$n_hole == 0)/nrow(cam_gap)), ")")` visits had complete CAM-ICU data.  Overall there are `r comma(sum(cam_gap$bt_days))` hospital days that should have CAM-ICU data, i.e., # of day in-between first and last CAM-ICU (including comma) for each visit.  The size of all holes in-between those days are  `r comma(sum(cam_gap$all_hole_size))` days, missing rate is `r percent(sum(cam_gap$all_hole_size)/as.numeric(sum(cam_gap$bt_days)))`.  Note that there are 11 days with conflicting CAM-ICUs assessed at exactly the same time.

```{r}

daily_status_f1 %>% 
  filter(status.today %in% c("Comatose", "Delirious", "Normal", "Unkown: conflicting CAM")) %>% 
  count(status.today) %>% 
  kable(caption = "Daily status in CAM-ICU visits") %>% 
    kable_styling(bootstrap_options = c("condensed", "bordered"), 
                full_width = F,
                position = "left") %>% 
  column_spec(1, bold = T)

daily_status_f1 %>% 
    filter(status.today %in% c( "Unkown: conflicting CAM"))
 
```



Here, we summarize status at visit level.  Note that A visit may have both coma and delirium.
```{r}
tibble(
  `Visit Status` = c("Total # of visits", 
           "# of visits with coma", "median # of days with coma (IQR)",
           "# of visits with delirium", "median # of days with delirium (IQR)",
           "# of visits with all normal", "# of visits with normal + unknown"),
  n = c(comma(nrow(visit_status2)), 
             comma(sum(visit_status2$any_coma)), paste0(coma_q[3], " (", coma_q[2], ", ", coma_q[4], ")"),
             comma(sum(visit_status2$any_del)), paste0(del_q[3], " (", del_q[2], ", ", del_q[4], ")"),
             comma(sum(visit_status2$all_norm)), comma(sum(visit_status2$norm_un)))
  ) %>% 
  kable(caption = "CAM-ICU visit summary") %>% 
    kable_styling(bootstrap_options = c("condensed", "bordered"), 
                full_width = F,
                position = "left") %>% 
  column_spec(1, bold = T) %>% 
  add_indent(c(3, 5))
```


## CAM-ICUs without admission/discharge records
```{r}
cam_noad <- cam_rass_daily %>% 
  filter(n != n_cam_na, n != n_cam_ua, n != n_cam_na + n_cam_ua) %>% 
  anti_join(
    daily_status_f1, by = c("grid", "dt")
  ) %>% 
  arrange(grid, dt) %>% 
  group_by(grid) %>% 
  mutate(date_diff = as.numeric(dt - lag(dt))) %>% 
  ungroup()
cam_noad$visit_id = NA
cam_noad$visit_id[1] = 1
for (i in 2:nrow(cam_noad)) {
  if (cam_noad$grid[i] != cam_noad$grid[i-1]) {
    cam_noad$visit_id[i] = 1
  } else if (cam_noad$grid[i] == cam_noad$grid[i-1] & cam_noad$dt[i] - cam_noad$dt[i-1] == 1) {
    cam_noad$visit_id[i] = cam_noad$visit_id[i-1]
  } else if (cam_noad$grid[i] == cam_noad$grid[i-1] & cam_noad$dt[i] - cam_noad$dt[i-1] > 1) {
    cam_noad$visit_id[i] = cam_noad$visit_id[i-1] + 1
  }
}
cam_noad_visit <- cam_noad %>% 
  group_by(grid, visit_id) %>% 
  summarise(first_cam_noad = min(dt),
            last_cam_noad = max(dt)) %>% 
  ungroup() 

cam_gap_noad <- sqldf::sqldf('SELECT *
        FROM cam_noad_visit as t1
        INNER JOIN cam_gap as t2 
        where t1.grid = t2.grid AND (first_cam_noad = dc_date + 1 OR last_cam_noad = adm_date - 1)') %>% 
  as_tibble() 
```

`r comma(length(unique(cam_noad$grid)))` patients had CAM-ICUs outside admission records. In total, it's `r comma(nrow(cam_noad))` days with CAM-ICU assessments (including coma), and were grouped into `r comma(nrow(cam_noad_visit))` visits.  
```{r}
cam_noad %>% 
  count(status.today)
```

If the visit's first or last CAM-ICU was right next to an exisiting discharge/admission records, it will be merged with that record.  And there are `r comma(nrow(cam_gap_noad))` such visits.
```{r}
cam_gap_noad %>% 
  select(grid:last_cam_noad, adm_id:last_cam)
```

Here's one example of CAM-ICU assessmens after recorded d/c date.  There are some gap between two blocks of CAM-ICUs.  Could this patient be out of ICU while in hospital?
```{r}
id = "R200860475"
cam_rass_daily %>% 
  filter(grid == id) %>% 
  select(grid, dt, status.today) 
cam_gap_noad %>% 
  filter(grid == id) %>% 
  select(grid:last_cam_noad, adm_id:last_cam)
```


There are also visit's whose first or last CAM-ICU were two days from an exisiting discharge/admission records

```{r}
id = "R200124788"
cam_rass_daily %>% 
  filter(grid == id) %>% 
  select(grid, dt, status.today)
cam_gap %>% 
  filter(grid == id) 
cam_noad %>% 
  filter(grid == id)  %>% 
  select(grid, visit_id, dt, status.today)
```


Due to lack of admission/discharge data, it's hard to know whether one day of missing CAM-ICU is a true hole or it's due to discharge from ICU/hospital.  Do we want to use RASS assessment to help with this?
```{r}
id = "R200376762"
cam_rass_daily %>% 
  filter(grid == id, year(dt) > 2007) %>% 
  select(grid, dt, status.today)
cam_noad %>% 
  filter(grid == id) %>% 
  select(grid, visit_id, dt, status.today)
```

Below is a patient who had two admission records that's apart only by one day.
```{r}
id = "R204461198"
cam_rass_daily %>% 
  filter(grid == id) %>% 
  select(grid, dt, status.today) 
cam_gap %>% 
  filter(grid == id) 
cam_noad %>% 
  filter(grid == id) %>% 
  select(grid, visit_id, dt, status.today) 
```

