---
title: "Identify Sepsis"
author: "Fei Ding"
date: "October 16, 2019"
output:
  html_document:
    df_print: paged
---




```{r setup, include = FALSE}
library(tidyverse)
library(scales)
library(kableExtra)
library(VennDiagram)
library(magrittr)

knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)
dob_dis <- read_csv("../output/dob_discrepancy.csv") 
cam_visits <- read_csv("../output/cam_stay_20190925.csv") 
sepsis3 <- read_csv("../output/sepsis3_20191014.csv")
rhee_sepsis <- read_csv("../output/sepsis_rhee_20191015.csv")
sepsis_comp <- cam_visits %>% 
  select(grid, adm_id) %>% 
  left_join(sepsis3 %>% 
              select(grid, adm_id, sofa, sepsis3, data_type)) %>% 
  left_join(rhee_sepsis %>% 
              select(grid, adm_id, rhee)) %>%  
  mutate(sepsis3_infection = if_else(is.na(sepsis3), 0, 1),
         sepsis3 = if_else(is.na(sepsis3), 0, sepsis3),
         rhee_infection = if_else(is.na(rhee), 0, 1),
         rhee = if_else(is.na(rhee), 0, rhee))

sepsis_comp %>% 
  semi_join(dob_dis)
sepsis_comp %>% 
  semi_join(dob_dis) %>% 
  distinct(grid)
```

# DOB discrepancy
As discussed in last meeting, the following 15 patients with DOB discrepancy > 10 days were removed from our final results.  
```{r}
dob_dis %>% filter(abs(diff) > 10) #24 pts total, 15 had diff > 10 
cam_visits %<>% anti_join(dob_dis, by = "grid")
sepsis3 %<>% anti_join(dob_dis, by = "grid")
rhee_sepsis %<>% anti_join(dob_dis, by = "grid")
sepsis_comp %<>% anti_join(dob_dis, by = "grid")
```


# Sepsis-3 definition

**NOTE:** Deviation from original definition were **bold (need more discussion)** or _italic (decided deviation)_.

1. For each CAM-ICU encounter, identify the first episode of suspected infection _within $\pm1$ day of hopital admission date_.
    * If the antibiotic was given first, the culture sampling must have been obtained within 1 day.
    * If the culture sampling was first, the antibiotic must have been ordered within 3 days.
    * The onset date of infection was the date on which the first of these two events occured.
    * __We used only blood culutre instead of body fluid cultures (blood, urine, cerebrospinal fluid, etc)__.
2. For each suspected infection, find worst daily SOFA score for the time window from 2 day before to 1 day after the onset date of infection.
    * Central nervous system SOFA score calculated _based on RASS Method C in the Modified SOFA score paper_.  
    * Respiration SOFA score calculated using both PaO~2~/FiO~2~(PF) ratio _and SpO~2~/FiO~2~(SF) ratio_.
        * FiO~2~ values > 100 were excluded. 1 < FiO~2~ <= 100 were converted to 0-1. 
        * SpO~2~ values > 100 were converted to 100.
    * __We used only creatinine data for renal SOFA, no urine output data__.
    * All the lab/drug dates are in the range of from 3 days before admission date to discharge date. 
    * All non-numerical lab values were converted to upper/lower limit or mean of the range.  
   
    
Rass score | PF or SF ratio and Ventilation |SOFA score
-----------|-----|------
$\ge$ 0 | PF $\ge$ 400 or SF $\ge$ 512 | 0  
-1 | PF < 400 or SF < 512 | 1
-2 | PF < 300 or SF < 357 | 2
-3 | PF < 200 or SF < 214 with ventilation | 3
-4 to -5 | PF < 100 or SF < 89 with ventilation | 4

3. Suspected infection with a SOFA score $\ge 2$ is considered sepsis __assuming baseline SOFA score is zero__.






## Distribution of SOFA score
`r paste0(comma(sum(sepsis3$sepsis3 == 1)), " (", percent(sum(sepsis3$sepsis3 == 1)/nrow(sepsis3)), ")")` out of the `r comma(nrow(sepsis3))` suspected infections had a SOFA score $\ge 2$.
```{r}
summary(sepsis3$sofa) 
ggplot(sepsis3) +
  geom_histogram(aes(x = sofa), binwidth = 1) +
  labs(title = "Distribution of SOFA score") 
```

## Missing Data
Missing data analysis below is based on the first worst SOFA of each suspected infection.  
Respiration system had the most missing data despite that we added SF ratio.  PaO2 had the most amount of data, which only covers fewer than 40% of infections.  SpO2 had FiO2 had even fewer data, covering fewer than 30% of infections.
```{r}
sepsis3 %>% 
  select(sofa_respiration, sofa_coagulation, sofa_liver, sofa_cardio, sofa_cns, sofa_renal) %>% 
  sapply(function(x) percent(sum(is.na(x))/length(x))) %>% 
  kable(caption = "Percent of missing data for each system", col.names = "% missing", full_width = F) %>% 
  kable_styling(bootstrap_options = "condensed", full_width = F)
sepsis3 %>% 
  count(data_type) %>% 
  mutate(pct = percent(n/nrow(sepsis3), accuracy = 0.01)) %>% 
  kable(caption = "Data composition of all suspected infections") %>% 
  kable_styling(bootstrap_options = "condensed", full_width = F)
```

Missing in SOFA < 2
```{r}
sepsis3 %>% 
  filter(sofa < 2) %>% 
  select(sofa_respiration, sofa_coagulation, sofa_liver, sofa_cardio, sofa_cns, sofa_renal) %>% 
  sapply(function(x) percent(sum(is.na(x))/length(x))) %>% 
  kable(caption = "Percent of missing data for each system", col.names = "% missing", full_width = F) %>% 
  kable_styling(bootstrap_options = "condensed", full_width = F)
sepsis3 %>% 
  filter(sofa < 2) %>% 
  count(data_type) %>% 
  mutate(pct = percent(n/sum(sepsis3$sofa < 2, na.rm = T), accuracy = 0.1)) %>% 
  kable(caption = "Data composition for SOFA < 2") %>% 
  kable_styling(bootstrap_options = "condensed", full_width = F)
```

# Rhee definition

1. For each CAM-ICU encounter, identify presumed serious infection .
    * Blood culture (BC) obtained _within $\pm1$ day of hopital admission date_, AND
    * $\ge 4$ QADs starting within $\pm2$ day of BC day.
        * __First QAD should be no later than discharge date, but no such limit for subsequent QADs__.
        * A gap of a single calendar day between the same antibiotic count as QADs.
        * If the paitent died before 4 QADs, the paitent can still qualify for presumed infection as long as they have consecutive QADs until 1 day prior to death. 
        * There must be at least one new IV/IM antibiotic within the $\pm2$ day window of BC day.
2. For each presumed infection, if any of the acute organ dysfunction criteria was met within the $\pm2$ day window of BC day, it is a sepsis.
    * Since we limited BC day within $\pm1$ day of hopital admission date, all of our infections are "present-on-admission infection".  Baseline lab values were the best value during hospitalization, from 3 days before admission date to discharge date.
    * __Our vasopressor don't have phenlyephrine or vasopressin__.
    * __We don't have eGFR data or ICD-9-CM code for end-stage kidney disease__.
    * __The platelet count criteria is <100 cells/$\mu$L, but I did not see any of our lab value lower than 0.1x10^3^/$\mu$L__.
        

# Compare two definitions
Out of `r comma(nrow(cam_visits))` ICU encounters with CAM-ICU assessments, 

* Using __Sepsis-3__ definition
    * `r paste0(comma(nrow(sepsis3)), " (", percent(nrow(sepsis3)/nrow(cam_visits)), ")")` had suspected infections
    * __`r paste0(comma(sum(sepsis3$sepsis3 == 1)), " (", percent(sum(sepsis3$sepsis3 == 1)/nrow(sepsis3)), ")")`__ of suspected infections had a SOFA score $\ge 2$.  
    * Overall, __`r percent(sum(sepsis3$sepsis3 == 1)/nrow(cam_visits))`__ ICU encounters had sepsis, that is __`r comma(sepsis3 %>% filter(sepsis3 == 1) %>% distinct(grid) %>% nrow())`__ distinct patients.  
* Using __Rhee__ definition
    * `r paste0(comma(nrow(rhee_sepsis)), " (", percent(nrow(rhee_sepsis)/nrow(cam_visits)), ")")` had presumed infections
    * __`r paste0(comma(sum(rhee_sepsis$rhee == 1)), " (", percent(sum(rhee_sepsis$rhee == 1)/nrow(rhee_sepsis)), ")")`__ of presumed infections had acute organ dysfunction.  
    * Overall, __`r percent(sum(rhee_sepsis$rhee == 1)/nrow(cam_visits))`__ ICU encounters had sepsis, that is __`r comma(rhee_sepsis %>% filter(rhee == 1) %>% distinct(grid) %>% nrow())`__ distinct patients.

```{r}
sensitivity <- sepsis_comp %>% filter(rhee == 1, sepsis3 == 1) %>% nrow() / 
  sepsis_comp %>% filter(sepsis3 == 1) %>% nrow()
specificity <- sepsis_comp %>% filter(rhee == 0, sepsis3 == 0) %>% nrow() / 
  sepsis_comp %>% filter(sepsis3 == 0) %>% nrow()
```

## Sepsis-3 vs. Rhee 
Sensitivity and specificity of Rhee definition are `r percent(sensitivity)` and `r percent(specificity)` respectively.

```{r}
sepsis_comp %<>% 
  mutate(sepsis3 = factor(sepsis3, levels = c(1, 0)),
         rhee = factor(rhee, levels = c(1, 0)),
         sepsis3_infection = factor(sepsis3_infection, levels = c(1, 0)),
         rhee_infection = factor(rhee_infection, levels = c(1, 0))
         )
xtabs(~ rhee + sepsis3, data = sepsis_comp, addNA = T)
grid.newpage()
venn.plot <- draw.pairwise.venn(area1      = sepsis_comp %>% filter(sepsis3 == 1) %>% nrow(),
                                area2      = sepsis_comp %>% filter(rhee == 1) %>% nrow(),
                                cross.area = sepsis_comp %>% filter(rhee == 1, sepsis3 == 1) %>% nrow(),
                                category   = c("Sepsis-3", "Rhee"),
                                fill = c("blue", "red"),
                                lty = "blank",
                                ext.pos = 180)
```

## Sepsis-3 infection vs. Rhee infection
```{r}
sensitivity <- sepsis_comp %>% filter(rhee_infection == 1, sepsis3_infection == 1) %>% nrow() / 
  sepsis_comp %>% filter(sepsis3_infection == 1) %>% nrow()
specificity <- sepsis_comp %>% filter(rhee_infection == 0, sepsis3_infection == 0) %>% nrow() / 
  sepsis_comp %>% filter(sepsis3_infection == 0) %>% nrow()
```


Sensitivity and specificity of Rhee definition of infection are `r percent(sensitivity)` and `r percent(specificity)` respectively.  In other words, all the infections identified by Rhee's definition are also infections under Sepsis-3 definition.
```{r}
xtabs(~ rhee_infection + sepsis3_infection, data = sepsis_comp, addNA = T)
grid.newpage()
venn.plot <- draw.pairwise.venn(area1      = sepsis_comp %>% filter(sepsis3_infection == 1) %>% nrow(),
                                area2      = sepsis_comp %>% filter(rhee_infection == 1) %>% nrow(),
                                cross.area = sepsis_comp %>% filter(rhee_infection == 1, sepsis3_infection == 1) %>% nrow(),
                                category   = c("Sepsis-3 Infection", "Rhee Infection"),
                                fill = c("blue", "red"),
                                lty = "blank",
                                cat.pos = c(0, 0),
                                cat.dist= c(-0.02, -0.02)
                                )

```

## Sepsis-3 vs. Rhee in infections identified by both definitions
```{r}
sensitivity <- sepsis_comp %>% filter(rhee_infection == 1, rhee == 1, sepsis3 == 1) %>% nrow() / 
  sepsis_comp %>% filter(rhee_infection == 1, sepsis3 == 1) %>% nrow()
specificity <- sepsis_comp %>% filter(rhee_infection == 1, rhee == 0, sepsis3 == 0) %>% nrow() / 
  sepsis_comp %>% filter(rhee_infection == 1, sepsis3 == 0) %>% nrow()
```
In these `r comma(sepsis_comp %>% filter(rhee_infection == 1) %>% nrow())` infections, Sensitivity and specificity of Rhee definition are `r percent(sensitivity)` and `r percent(specificity)` respectively.  
So even when we only consider infections identified by both definitions, Sepsis-3 still identify more sepsis compared to Rhee. 
```{r}
sepsis_comp %>% 
  filter(rhee_infection == 1) %>% 
  with(xtabs(~rhee+sepsis3))
grid.newpage()
venn.plot <- draw.pairwise.venn(area1      = sepsis_comp %>% filter(rhee_infection == 1, sepsis3 == 1) %>% nrow(),
                                area2      = sepsis_comp %>% filter(rhee_infection == 1, rhee == 1) %>% nrow(),
                                cross.area = sepsis_comp %>% filter(rhee_infection == 1, sepsis3 == 1, rhee == 1) %>% nrow(),
                                category   = c("Sepsis-3", "Rhee"),
                                fill = c("blue", "red"),
                                lty = "blank",
                                ext.pos = 180
)

```

