---
title: "Data of updated GRIDs"
author: "Fei Ding"
date: "August 21, 2019"
output:
  html_document:
    df_print: paged
---


```{r setup, include = FALSE}
library(tidyverse)
library(readxl)
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)

load("../output/data_raw.RData")
changed_grid <- read_excel("../../Mito Delirium BioVU Data/Data/Changed_GRIDS.xlsx")
Hmisc::describe(changed_grid)
changed_grid <- distinct(changed_grid)
names(changed_grid) <- tolower(names(changed_grid))
Hmisc::describe(changed_grid)

## see how many duplicated data
length(unique(static_raw$grid))
sum(unique(static_raw$grid) %in% unique(changed_grid$old_grid))
sum(unique(static_raw$grid) %in% unique(changed_grid$updated_grid))

length(unique(discharge_raw$grid))
sum(discharge_raw$grid %in% changed_grid$updated_grid)
sum(unique(discharge_raw$grid) %in% unique(changed_grid$old_grid))
sum(unique(discharge_raw$grid) %in% changed_grid$updated_grid)

length(unique(cam_raw$grid))
sum(unique(cam_raw$grid) %in% changed_grid$old_grid)
sum(unique(cam_raw$grid) %in% unique(changed_grid$updated_grid))

sum(unique(discharge_raw$grid) %in% cam_raw$grid)
```

## Old GRID in our data, updated GRID NOT in our data
```{r}
## old GRID in our data, but updated GRID not in our data
changed_grid %>% 
  filter(old_grid %in% static_raw$grid, ! updated_grid %in% static_raw$grid) %>% 
  arrange(old_grid) 
```

### Example 1
Data pulled using old GRID
```{r}
discharge_raw %>% 
  filter(grid == "R200002089")
cam_raw %>% 
  filter(grid == "R200002089")
```
No Data pulled using corresponding updated GRID
```{r, echo=T}
discharge_raw %>% 
  filter(grid == "R215108619")
cam_raw %>% 
  filter(grid == "R215108619")
```

### Example 2
Data pulled using old GRID
```{r}
discharge_raw %>% 
  filter(grid == "R200302879")
cam_raw %>% 
  filter(grid == "R200302879")
```
No data pulled using corresponding updated GRID
```{r, echo=T}
discharge_raw %>% 
  filter(grid == "R270390013")
cam_raw %>% 
  filter(grid == "R270390013")
```

## Both old and updated GRIDs in our data
```{r}
changed_grid %>% 
  filter(old_grid %in% static_raw$grid, updated_grid %in% static_raw$grid) %>% 
  arrange(old_grid)
```

Different data pulled using corresponding old/updated GRIDs

### Example 3
3 admission records vs. 2 admission records
```{r}
discharge_raw %>% 
  filter(grid == "R200065814")
discharge_raw %>% 
  filter(grid == "R235732375")
```
11 CAM-ICU recods vs. 1 CAM-ICU records
```{r}
cam_raw %>% 
  filter(grid == "R200065814")
cam_raw %>% 
  filter(grid == "R235732375")
```
### Example 4
8-day hospital stay vs. 2-day hospital stay
```{r}
discharge_raw %>% 
  filter(grid == "R200189675")
discharge_raw %>% 
  filter(grid == "R239370958")
```
91 CAM-ICU records vs. 9 CAm-ICU records
```{r}
cam_raw %>% 
  filter(grid == "R200189675")
cam_raw %>% 
  filter(grid == "R239370958")
```

