---
title: "BioVU daily status"
author: "Fei Ding"
date: "March 19, 2019"
output:
  html_document:
    toc: true
    theme: united
---

```{r setup, include = FALSE}
library(tidyverse)
library(scales)
library(lubridate)
library(kableExtra)

load("../output/data_raw.RData")

rm(static_raw)
rm(discharge_raw)

cam_raw <- cam_raw %>% 
  rename(cam_time = cam_date) %>% 
  mutate(dttm = mdy_hms(cam_time),
         cam_date = as_date(dttm)) %>% 
  arrange(grid, dttm)
rass_raw <- rass_raw %>% 
  rename(rass_time = rass_date) %>% 
  mutate(dttm = mdy_hms(rass_time),
         rass_date = as_date(dttm)) %>% 
  arrange(grid, dttm) 
cam_icu <- cam_raw %>% 
  filter(cam_test_name == "CAM-ICU") %>% 
  select(grid, cam_date, dttm, cam_value)
cam_dup_row <- cam_icu %>% 
  select(grid, dttm) %>% 
  duplicated() %>% 
  which()
cam_icu_dup <- cam_icu %>% 
  slice(c(cam_dup_row, cam_dup_row - 1)) %>%
  arrange(grid, dttm, cam_value) 
cam_icu_dup_w <- cam_icu_dup %>% 
  group_by(grid, dttm) %>% 
  mutate(cam_value.x = lag(cam_value)) %>% 
  ungroup() %>% 
  filter(!is.na(cam_value.x)) %>% 
  rename(cam_value.y = cam_value)
rass_dup_row <- rass_raw %>% 
  select(grid, dttm) %>% 
  duplicated() %>% 
  which()
rass_dup <- rass_raw %>% 
  select(-rass_time) %>% 
  slice(c(rass_dup_row, rass_dup_row - 1)) %>% 
  arrange(grid, dttm, rass_score)
rass_dup_w <- rass_dup %>% 
  group_by(grid, dttm) %>% 
  mutate(rass_score.x = lag(rass_score),
         rass_score_test_name.x = lag(rass_score_test_name)) %>% 
  ungroup() %>% 
  filter(!is.na(rass_score.x)) %>% 
  rename(rass_score.y = rass_score,
         rass_score_test_name.y = rass_score_test_name)
cam_rass_dup <-  cam_icu_dup_w %>% 
  select(grid, dttm, cam_date, cam_value.x, cam_value.y) %>% 
  semi_join(
    rass_dup %>% 
      distinct(grid, dttm) %>% 
      dplyr::intersect(
        cam_icu_dup %>% 
          distinct(grid, dttm) 
      ),
    by = c("grid", "dttm")
  ) %>% 
  left_join(
    rass_dup_w %>% 
      select(grid, dttm, rass_score.y, rass_score.x),
    by = c("grid", "dttm")
  )
rass_dup_red <- rass_dup_w %>% 
  select(-rass_score.y, -rass_score_test_name.y) %>% 
  rename(rass_score = rass_score.x,
         rass_score_test_name = rass_score_test_name.x) 
rass_clean <- rass_raw %>% 
  select(-rass_time) %>% 
  anti_join(rass_dup_red, by = c("grid", "dttm")) %>% 
  bind_rows(rass_dup_red)
cam_icu_dup_rass <- cam_icu_dup_w %>%
  left_join(rass_clean, 
            by = c("grid", "dttm")) %>% 
  select(grid, dttm, cam_date, cam_value.x, cam_value.y, rass_date, rass_score, rass_score_test_name)
cam_icu_dup_rass <- cam_icu_dup_rass %>% 
  mutate(cam_value = case_when(
    cam_value.x == "Delirium present" & cam_value.y == "No delirium" ~ "Unk",
    cam_value.x == "Negtiv" & cam_value.y == "Postiv" ~ "Unk",
    cam_value.x %in% c("Negtiv", "Postiv") & cam_value.y == "Unases" & rass_score %in% -5:-4 ~ "UA",
    cam_value.x %in% c("Negtiv", "Postiv") & cam_value.y == "Unases" & !rass_score %in% -5:-4 ~ cam_value.x
  )) 
cam_icu_clean <- cam_icu %>% 
  select(grid, cam_value, dttm, cam_date) %>% 
  anti_join(cam_icu_dup_rass, by = c("grid", "dttm")) %>% 
  bind_rows(select(cam_icu_dup_rass, grid, cam_value, dttm, cam_date))
cam_rass <- cam_icu_clean %>% 
  full_join(select(rass_clean, -rass_score_test_name), 
            by = c("grid", "dttm")) %>% 
  select(grid, dttm, cam_date, cam_value, rass_date, rass_score) %>% 
  arrange(grid, dttm) %>% 
  mutate(
    dt = as_date(dttm),
    rass = case_when(
      rass_score %in% -5:4 ~ rass_score
    ),
    cam = case_when(
      rass_score %in% -5:-4 ~ "UA",
      cam_value %in% c("Delirium present", "Postiv") ~ "Positive",
      cam_value %in% c("No delirium", "Negtiv") ~ "Negative"
    )
  )

cam_rass_daily <- cam_rass %>% 
  group_by(grid, dt) %>% 
  count(cam) %>% 
  ungroup() %>% 
  spread(key = cam, value = n, fill = 0L) %>% 
  bind_cols(
    cam_rass %>% 
      mutate(
        rass_c = case_when(
          rass %in% -5:-4 ~ "n_rass_coma",
          rass %in% -3:4 ~ "n_rass_nocoma",
          is.na(rass) ~ "n_rass_na"
        )
      ) %>% 
      group_by(grid, dt) %>% 
      count(rass_c) %>% 
      ungroup() %>% 
      spread(key = rass_c, value = n, fill = 0L) 
  ) %>% 
  rename(n_cam_pos = Positive,
         n_cam_neg = Negative,
         n_cam_ua = UA,
         n_cam_na = `<NA>`) %>% 
  mutate(n = n_rass_coma + n_rass_na + n_rass_nocoma) %>% 
  select(-grid1, -dt1)

cam_rass_daily <- cam_rass_daily %>% 
  mutate(
    status.today = case_when(
      n_cam_pos > 0 ~ "Delirious",
      n_rass_coma > 0 ~ "Comatose",
      n_cam_neg > 0 ~ "Normal",
      n == n_cam_na & n_rass_nocoma > 0 ~ "Unknown: RASS only",
      n == n_cam_na & n == n_rass_na ~ "Unknown: No CAM nor RASS"
    )) 

```

Only *CAM-ICU* data was used.  *Unable to Assess CAM-ICU* data was checked separately.  They either did not provide any new information in addition to RASS score or having conflict RASS values.  So they were not included here.

# Data Management before merge
In order to merge RASS data and CAM-ICU data by assessment time, duplicated CAM-ICU or RASS need to be reduced first.

## Duplicated assessment time in RASS
285 pairs of RASS scores were measured at duplicated assessment time points.

Current approach: 
Take the first column, the lower value of RASS score so that if one of the duplicated value is -5 or -4, it was counted.

Should we take another approach? Say, consider nearby RASS score?  CAM-ICU values are not very helpful here.
```{r, echo = FALSE}
rass_dup_w %>%
  count(rass_score.x, rass_score.y) %>%
  kable(caption = "Duplicated RASS scores") %>% 
  kable_styling(bootstrap_options = c("condensed", "bordered"), 
                full_width = F,
                position = "left") %>% 
  row_spec(c(2:5,7:11), background = "#F0E442", color = "black") %>%
  row_spec(c(6, 22, 26), background = "grey", color = "white") %>% 
  column_spec(1, bold = T)
```

## Duplicated assessment time in CAM-ICU
92 pairs of CAM-ICU values were measured at duplicated assessment time points.
```{r, echo = FALSE}
cam_icu_dup_w %>% 
  group_by(cam_value.x, cam_value.y) %>% 
  count() %>% 
  kable(caption = "Duplicated CAM-ICU values") %>% 
  kable_styling(bootstrap_options = c("condensed", "bordered"), 
                full_width = F,
                position = "left")
```

CAM-ICU duplication was resolved based on the corresponding RASS score.

* If RASS score was -5 or -4, CAM-ICU was assigned *UA* (unassessible)
* If RASS score was not -5 or -4, and one of the duplicated CAM-ICU was *Unases*, the other CAM-ICU value (*Postive* or *Negtive*) was taken.
* If duplicated CAM-ICUs were positive and negative, CAM-ICU was assigned *Unk* (Unknown)
```{r, echo = FALSE}
cam_icu_dup_rass %>% 
  count(cam_value.x, cam_value.y, rass_score, cam_value) %>% 
   mutate(
    cam_value = cell_spec(
      cam_value, "html",
      color = "white",
      background  = case_when(
        cam_value == "Postiv" ~ "#D55E00", 
        cam_value == "Negtiv" ~ "#56B4E9",
        cam_value == "Unk" ~ "black",
        cam_value == "UA" ~ "gray")
      )
    ) %>% 
  kable(format = "html", escape = F, caption = "Reduce Duplicated CAM") %>% 
  kable_styling(bootstrap_options = c("bordered", "condensed"), 
                full_width = F,
                position = "left")
```


### Duplication in both RASS and CAM-ICU
10 of the assessment time points had both duplicated CAM-ICU values and RASS scores. Based on the reduction rules above, they were reduced as following:
```{r, echo = FALSE}
cam_icu_dup_rass %>% 
  select(grid, dttm, cam_value) %>% 
  inner_join(cam_rass_dup, by = c("grid", "dttm")) %>% 
  select(grid, dttm, cam_value.y, cam_value.x, rass_score.x, rass_score.y, cam_value) %>% 
  arrange(cam_value.y) %>% 
  mutate(
    cam_value = cell_spec(
      cam_value, "html",
      color = "white",
      background  = case_when(
        cam_value == "Postiv" ~ "#D55E00", 
        cam_value == "Negtiv" ~ "#56B4E9",
        cam_value == "UA" ~ "black",
        cam_value == "Unk" ~ "gray")
      )
    ) %>% 
  kable(format = "html", escape = F, caption = "10 time points with both duplicated CAM-ICU and RASS") %>% 
  kable_styling(bootstrap_options = c("bordered", "condensed"), 
                full_width = F,
                position = "left")
```

# Merge CAM and RASS to get daily status
RASS data and CAM-ICU data were merged by assessment time.  All the RASS and CAM-ICU assessments were kept.

## Assessment level
For each assessment, the following rules were applied in order:

* RASS values not in the range [-5, 4] were assigned to *<NA>*
* If RASS value is -5 or -4, CAM-ICU was assinged to *UA*
* CAM-ICU values *Delirium present* and *Postive* were assinged to *Positive*
* CAM-ICU values *No delirium* and *Negativ* were assigned to *Negative*
* CAM-ICU values *Unases* and *Unk* was assgined to *<NA>*



```{r, echo = FALSE}
xtabs( ~ rass_score + cam_value, data = cam_rass, addNA = T) %>% 
  kable(caption = "CAM-ICU vs RASS before cleaning up") %>% 
  kable_styling(bootstrap_options = c("condensed", "bordered"), 
                full_width = F,
                position = "left") %>% 
  column_spec(7:9, background = "grey", color = "white") %>%
   column_spec(c(2,5), background = "#D55E00", color = "white") %>%
   column_spec(3:4, background = "#56B4E9", color = "white") %>%
  row_spec(3:4, background = "black", color = "white") %>% 
  column_spec(1, bold = T)
xtabs( ~ rass + cam, data = cam_rass, addNA = T) %>% 
  kable(caption = "CAM-ICU vs RASS after cleaning up") %>% 
   kable_styling(bootstrap_options = c("bordered", "condensed"), 
                full_width = F,
                position = "left") %>% 
  add_header_above(c("RASS score" = 1, "CAM-ICU value" = length(unique(cam_rass$cam)))) %>% 
  column_spec(1, bold = T) %>% 
  column_spec(2, background = "#56B4E9", color = "white") %>% 
  column_spec(3, background = "#D55E00", color = "white") %>%
  column_spec(4, background = "black", color = "white") %>%
  column_spec(5, background = "grey", color = "white") 
  
```

## Daily status
To get daily status, the following rules were applied to each calendar day:

* *Delirious* if any CAM-ICU value was Positive
* Otherwise *comatose* if any RASS value was -5 or -4
* Otherwise *Normal* if any CAM-ICU value was Negative
* Otherwise *Unknown: RASS only* if all CAM-ICU values were missing and at least one non-missing RASS value
* Otherwise *Unknown: No CAM nor RASS* if all CAM-ICU and RASS values were missing


There are total `r comma(nrow(cam_rass_daily))` person-days.
```{r, echo = FALSE}
cam_rass_daily %>% 
  count(status.today) %>% 
  mutate(pct = percent(nn/nrow(cam_rass_daily))) %>% 
  kable(caption = "Daily Status") %>% 
  kable_styling(bootstrap_options = c("condensed", "bordered"), 
                full_width = F,
                position = "left") %>% 
  column_spec(1, bold = T)
  
```


## Unknown: RASS only category
For the last category *Unknown: RASS only*, 20,556 person-days were in CAM-ICU file but CAM-ICU value was missing.
668,793 person-days not in CAM-ICU file. of which 16,324 were in CAM file with some CAM values other than CAM-ICU, the other 652,469 person-days were not found in CAM at all.   This is not surprising given that RASS file has almost twice number of person-days compared to CAM file.

See a few examples

The first example showed a case where for each day there was only one RASS assessment but no CAM-ICU.

The second example showed a case where there were several RASS assessment but no CAM-ICU during a day.
```{r, echo = FALSE}
## see a few examples
cam_rass_daily %>% 
  filter(grid == "R200000649", dt %within% interval("2011-06-20", "2011-06-26")) %>% 
  select(grid, dt, status.today) %>% 
   mutate(
    status.today = cell_spec(
      status.today, "html",
      color = "white",
      background  = case_when(
        status.today == "Delirious" ~ "#D55E00", 
        status.today == "Normal" ~ "#56B4E9",
        status.today == "Comatose" ~ "black",
        status.today == "Unknown: RASS only" ~ "gray")
      )
    ) %>% 
   kable(format = "html", escape = F, caption = "Example 1, daily level") %>% 
  kable_styling(bootstrap_options = c("condensed", "bordered"), 
                full_width = F,
                position = "left") 
cam_rass %>% 
  filter(grid == "R200000649", dt %within% interval("2011-06-20", "2011-06-26")) %>% 
  select(grid,  dttm, cam_value, rass_score, cam) %>% 
   mutate(
    cam = cell_spec(
      cam, "html",
      color = "white",
      background  = case_when(
        cam == "Positive" ~ "#D55E00", 
        cam == "Negative" ~ "#56B4E9",
        cam == "UA" ~ "black",
        is.na(cam) ~ "gray")
      )
    ) %>% 
  kable(format = "html", escape = F, caption = "Example 1, assessment level") %>% 
  kable_styling(bootstrap_options = c("bordered", "condensed"), 
                full_width = F,
                position = "left")
  
cam_rass_daily %>% 
  filter(grid == "R200002089", dt %within% interval("2007-08-25", "2007-08-31")) %>% 
  select(grid, dt, status.today) %>% 
    mutate(
    status.today = cell_spec(
      status.today, "html",
      color = "white",
      background  = case_when(
        status.today == "Delirious" ~ "#D55E00", 
        status.today == "Normal" ~ "#56B4E9",
        status.today == "Comatose" ~ "black",
        status.today == "Unknown: RASS only" ~ "gray")
      )
    ) %>% 
   kable(format = "html", escape = F, caption = "Example 2, daily level") %>% 
  kable_styling(bootstrap_options = c("condensed", "bordered"), 
                full_width = F,
                position = "left") 
cam_rass %>% 
  filter(grid == "R200002089", dt %within% interval("2007-08-25", "2007-08-31")) %>% 
  select(grid,  dttm, cam_value, rass_score, cam) %>% 
  mutate(
    cam = cell_spec(
      cam, "html",
      color = "white",
      background  = case_when(
        cam == "Positive" ~ "#D55E00", 
        cam == "Negative" ~ "#56B4E9",
        cam == "UA" ~ "black",
        is.na(cam) ~ "gray")
      )
    ) %>% 
  kable(format = "html", escape = F, caption = "Example 1, assessment level") %>% 
  kable_styling(bootstrap_options = c("bordered", "condensed"), 
                full_width = F,
                position = "left")
```


# Questions to discuss

* Do we want to first make sure all dates are within admission/discharge range? Are RASS assessed outside ICU?
* Do we want to do something similar to before, assign a CAM-ICU value based on RASS score when CAM-ICU is missing? 
* Last time, Dr. Tang proposed for days with only negative CAM-ICU, we may also want to impute a CAM-ICU value, which will add more imputation.  Do we want to do this?  I think if CAM-ICU was assessed often enough (on the CAM-ICU sheet I found, it says every 8-12 hours), maybe we can just accept the negative value.
* Regarding imputation, we will impute at daily level for sure. However, since daily status is a combination of CAM-ICU and RASS value, if we want to use CAM-ICU and RASS at individual assessment to impute, isn't this kind of at assessment level?