---
title: "BioVU CAM-ICU data"
author: "Fei Ding"
date: "Feburary 28, 2019"
output:
  html_document
---

 

```{r setup, include = FALSE}
options(width = 200)
library(tidyverse)
library(Hmisc)
library(scales)
library(lubridate)
library(kableExtra)
load("../output/data_raw.RData")

static_raw <- as.tibble(static_raw)
discharge_raw <- as.tibble(discharge_raw)
cam_raw <- as.tibble(cam_raw)
rass_raw <- as.tibble(rass_raw)
cam_icu_daily_ct <- read_csv("../output/cam_icu_daily_summary.csv")

rass_raw <- rass_raw %>% 
  rename(rass_time = rass_date) %>% 
  mutate(rass_dttm = mdy_hms(rass_time),
         rass_date = as_date(rass_dttm)) %>% 
  arrange(grid, rass_dttm) 
cam_raw <- cam_raw %>% 
  distinct() %>% 
  rename(cam_time = cam_date) %>% 
  mutate(cam_dttm = mdy_hms(cam_time),
         cam_date = as_date(cam_dttm)) %>% 
  arrange(grid, cam_dttm)
```

## Number of calendar days without any CAM-ICU assessments
For each CAM-ICU assessment day, total number of CAM-ICU assessments, number of negative, postive, unassessed, and missing CAM-ICU were calculated.   It is an interesting summary to look at.
The example below shows that the patient had all negative CAM-ICU in the first five days, then had a few days with either positive or unassessed CAM-ICUs, then several days with both positive and negative CAM-ICU, and at last, go back to all negative CAM-ICU.  
```{r, echo = F}
cam_icu_daily_ct %>% 
  filter(grid == "R200002089") %>% 
  select(-composition) %>% 
  print(n = 50)
```

Each calendar day was categorized based on assessment result composition.
out of the `r nrow(cam_icu_daily_ct)` person-days, `r sum(cam_icu_daily_ct$composition %in% c("All Missing", "All Unassessed", "Missing + Unassessed"))` person-day had all CAM-ICU unassessed or missing
```{r, echo = F}
count(cam_icu_daily_ct, composition) %>% 
  print(n = 10)
```

For the `r sum(cam_icu_daily_ct$composition %in% c("All Missing", "All Unassessed", "Missing + Unassessed"))` person-days with all CAM-ICU unassessed or missing, we also checked whether there are any CAM measurements other than CAM-ICU on that day.
22 patient-days have some other CAM measure (mostly no Delirium, on Yes/CAM Total = 6), often paired *CAM Total* and *Delirium Present*, sometimes just one.
```{r, echo = F}
cam_icu_daily_ct %>% 
  filter(composition %in% c("All Missing", "All Unassessed", "Missing + Unassessed")) %>% 
  left_join(
    cam_raw, by = c("grid", "cam_date")
  ) %>% 
  filter(cam_test_name != "CAM-ICU") %>% 
  select(grid, cam_date, n, composition, cam_dttm, cam_test_name, cam_value) %>% 
  print(n = 100)
```


## See a few examples of CAM-RASS conflicts
CAM-ICU and RASS were merged by assessment time to find CAM-RASS conflicts.
```{r, include = F}
cam_icu <- cam_icu <- cam_raw %>% 
  filter(cam_test_name == "CAM-ICU") %>% 
  select(-cam_test_name, -cam_time) %>% 
  mutate(cam_value = ifelse(is.na(cam_value), "Missing", cam_value))
cam_rass <- cam_icu %>% 
  full_join(select(rass_raw, -rass_score_test_name, -rass_time), 
            by = c("grid", "cam_dttm" = "rass_dttm")) %>% 
  rename(dttm = cam_dttm) %>% 
  select(grid, dttm, cam_date, cam_value, rass_date, rass_score) %>% 
  arrange(grid, dttm) %>% 
  mutate(
    conflict = case_when(
      rass_score %in% -3:4 & cam_value %in% c("Unases", "Missing") ~ "Type II",
      rass_score %in% -5:-4 & cam_value %in% c("Postiv", "Delirium present", "Negtiv", "No delirium") ~ "Type I"
    )
  ) 
```

There are two types of conflicts. Type I conflict: RASS is -5 or -4, but CAM was either positive or negative; Type II conflict: RASS is -3 to 4, and CAM was unassessed.  They are highlighted in grey and yellow respectively in the tables below.  There are way more Type II conflicts  (`r sum(cam_rass$conflict %in% "Type II")`) compared to Type I conflicts (`r sum(cam_rass$conflict %in% "Type I")`).  Type II conflict may be ok, we just need to make sure we call coma based on RASS value, not unassessed CAM. 
```{r, echo = F}
cam_rass %>% 
  filter(rass_score %in% -5:4 | is.na(rass_score)) %>% 
  with(xtabs( ~ rass_score + cam_value, addNA = T)) %>% 
  kable(caption = "CAM-ICU vs. RASS, focus on RASS -5 to 4") %>% 
  kable_styling(bootstrap_options = c("condensed", "bordered"), 
                full_width = F,
                position = "left") %>% 
  add_header_above(c("RASS score" = 1, "CAM-ICU value" = length(unique(cam_rass$cam_value)))) %>% 
  row_spec(1:2, background = "grey", color = "white") %>% 
  column_spec(c(3, 7), background = "#F0E442", color = "black")
 

conf_list <- cam_rass %>% 
  filter(grid == "R200002089", 
         as_date(dttm) %within% interval("2007-08-22", "2007-09-01")  ) %>% 
  arrange(grid, dttm)
conf_list %>% 
   select(grid, dttm, cam_value, rass_score) %>% 
  mutate(
    cam_value = cell_spec(
      cam_value, "html",
      color = "white",
      background  = case_when(
        cam_value %in% c("Postiv", "Delirium present") ~ "#D55E00", 
        cam_value %in% c("Negtiv", "No delirium") ~ "#56B4E9",
        T ~ "black")
      )
    ) %>% 
  kable(format = "html", escape = F, caption = "Example 1") %>% 
  kable_styling(bootstrap_options = c("bordered", "condensed"), 
                full_width = F,
                position = "left") %>% 
  row_spec(with(conf_list, which(conflict %in% "Type I")), background = "grey", color = "white") %>% 
  row_spec(with(conf_list, which(conflict %in% "Type II")), background = "#F0E442", color = "black")

conf_list <- cam_rass %>% 
  filter(grid == "R202019861", 
         as_date(dttm) %within% interval("2012-03-28", "2012-03-31")  ) %>% 
  arrange(grid, dttm)
conf_list %>% 
  select(grid, dttm, cam_value, rass_score) %>% 
  mutate(
    cam_value = cell_spec(
      cam_value, "html",
      color = "white",
      background  = case_when(
        cam_value %in% c("Postiv", "Delirium present") ~ "#D55E00", 
        cam_value %in% c("Negtiv", "No delirium") ~ "#56B4E9",
        T ~ "black")
      )
    ) %>% 
  kable(format = "html", escape = F, caption = "Example 2") %>% 
  kable_styling(bootstrap_options = c("bordered", "condensed"), 
                full_width = F,
                position = "left") %>% 
  row_spec(with(conf_list, which(conflict %in% "Type I")), background = "grey", color = "white") %>% 
  row_spec(with(conf_list, which(conflict %in% "Type II")), background = "#F0E442", color = "black")

conf_list <- cam_rass %>% 
  filter(grid == "R201753181", 
         as_date(dttm) %within% interval("2010-06-27", "2010-06-29")  ) %>% 
  arrange(grid, dttm)
conf_list %>% 
  select(grid, dttm, cam_value, rass_score) %>% 
  mutate(
    cam_value = cell_spec(
      cam_value, "html",
      color = "white",
      background  = case_when(
        cam_value %in% c("Postiv", "Delirium present") ~ "#D55E00", 
        cam_value %in% c("Negtiv", "No delirium") ~ "#56B4E9",
        T ~ "black")
      )
    ) %>% 
  kable(format = "html", escape = F, caption = "Example 3") %>% 
  kable_styling(bootstrap_options = c("bordered", "condensed"), 
                full_width = F,
                position = "left") %>% 
  row_spec(with(conf_list, which(conflict %in% "Type I")), background = "grey", color = "white") %>% 
  row_spec(with(conf_list, which(conflict %in% "Type II")), background = "#F0E442", color = "black")
```

Of the `r sum(cam_rass$conflict %in% "Type II")` Type I conflicts, 679 have another CAM test *Unable to assess CAM-ICU* recorded at the same time point. Below is an example.
```{r, echo = F}
ex1 <- cam_rass %>% 
  filter(grid == "R201908155", as_date(dttm) %within% interval("2016-01-02", "2016-01-04")) %>% 
  select(grid, dttm, cam_value, rass_score) 
ex2 <- cam_raw %>% 
  filter(grid == "R201908155", cam_test_name != "CAM-ICU", cam_date %within% interval("2016-01-02", "2016-01-04")) %>% 
  select(grid, cam_dttm, cam_value)
ex2 %>% 
  full_join(ex1, by = c("grid", "cam_dttm" = "dttm")) %>% 
  rename(`CAM-ICU` = cam_value.y,
         `Unable to assess CAM-ICU` = cam_value.x) %>% 
  arrange(grid, cam_dttm) %>% 
  print(n = 100)

```

### year distribution of CAM-RASS conlict
```{r, echo = F}
addmargins(xtabs( ~ year(dttm) + conflict, cam_rass, addNA = T)) %>% 
  kable(col.names = c("Type I", "Type II", "No conflict", "All")) 
```

## Duplicated CAM-ICU
There are 92 pairs of duplicated CAM-ICUs.  Some of the assessments also had the CAM-ICU conflict problem.
```{r, echo = F}
cam_icu_dup <- cam_icu %>% 
  filter(duplicated(select(cam_icu, grid, cam_dttm), fromLast = T)|
           duplicated(select(cam_icu, grid, cam_dttm))) %>% 
  arrange(grid, cam_dttm, cam_value)

cam_icu_dup %>% 
  select(grid, cam_value, cam_dttm) %>% 
  filter(duplicated(select(cam_icu_dup, grid, cam_dttm), fromLast = T)) %>% 
  inner_join(
    cam_icu_dup %>% 
      select(grid, cam_value, cam_dttm) %>% 
      filter(duplicated(select(cam_icu_dup, grid, cam_dttm))),
    by = c("grid", "cam_dttm")
  ) %>% 
  group_by(cam_value.x, cam_value.y) %>% 
  count() %>% 
  print(n = 10)

cam_icu_dup %>% 
  select(grid, cam_value, cam_dttm) %>% 
  filter(duplicated(select(cam_icu_dup, grid, cam_dttm), fromLast = T)) %>% 
  inner_join(
    cam_icu_dup %>% 
      select(grid, cam_value, cam_dttm) %>% 
      filter(duplicated(select(cam_icu_dup, grid, cam_dttm))),
    by = c("grid", "cam_dttm")
  ) %>% 
  inner_join(filter(cam_rass, !is.na(conflict)), by = c("grid", "cam_dttm" = "dttm")) %>% 
  arrange(grid, cam_dttm) %>% 
  select(grid, cam_dttm, cam_value, rass_score, conflict) %>% 
  print(n = 100)

```

